### 2014-11-14 conclusion.tex
first draft conclusion

In this paper, we have proposed a method for generating 3D CAD model
exemplar detector templates on-the-fly, based on a novel variant of
WHO features called NZ-WHO. As it circumvents the need for
calibration, it is computationally efficient, allowing it to be run in
an on-the-fly setting. As a result, we can use our method to enrich
existing object class detections with additional information such as
precise 3D pose and 3D CAD model exemplar.